---
title: "Coding_Challange_7"
author: "Arpan"
date: "2025-04-02"
output: 
  md_document:
    variant: gfm
  html_document:
    toc: true
    toc_float: true
---

# Question 1

**Read in the data called “PlantEmergence.csv” using a relative file path and load the following libraries. tidyverse, lme4, emmeans, multcomp, and multcompView. Turn the Treatment , DaysAfterPlanting  and Rep into factors using the function as.factor**
**STAND$Treatment <- as.factor(STAND$Treatment) # example shown here.**

```{r Question 1}
#Loading libraries
library(tidyverse)
library(emmeans)
library(multcomp)
library(multcompView)
library(lme4)

plantdata<-read.csv("LinearModel/PlantEmergence.csv", na.strings="na") #Loading the plant emergence data into R

#Converting treatment, days after lanting and replication into factors
plantdata$Treatment <- as.factor(plantdata$Treatment) 
plantdata$DaysAfterPlanting<-as.factor(plantdata$DaysAfterPlanting)
plantdata$Rep<-as.factor(plantdata$Rep)
```

# Question 2

2. Fit a linear model to predict Emergence using Treatment and DaysAfterPlanting along with the interaction. Provide the summary of the linear model and ANOVA results. 

```{r Q3:Linear model with interaction}
lm1<-lm(Emergence~Treatment*DaysAfterPlanting,data=plantdata)
summary(lm1)
anova(lm1)
```

# Question 3

Based on the results of the linear model in question 2, do you need to fit the interaction term? Provide a simplified linear model without the interaction term but still testing both main effects. Provide the summary and ANOVA results. Then, interpret the intercept and the coefficient for Treatment 2.

Based on the results of the linear model in question 2, we dont need to fit the interaction term because the interaction is not significant as shown by anova and summary of the model.

```{r}
lm2<-lm(Emergence~Treatment+DaysAfterPlanting,data=plantdata)
summary(lm2)
anova(lm2)
confint(lm2)
```

We found that Treatment 2 had 134.53 (+-6.775364;+-95% CI)  units less emergence time than Treatment 1.
The intercept is the average emergence days for treatment 1 7 days after planting which is 182.163 days.
The coefficient of Treatment 2 is the difference in emergence days between treatment 2 and treatment 1.

# Question 4
**Calculate the least square means for Treatment using the emmeans package and perform a Tukey separation with the compact letter display using the cld function. Interpret the results.**

```{r emmeams}
lsmeans<-emmeans(lm2,~Treatment)#|DaysAfterPlanting) 
results_lsmeans<-cld(lsmeans,alpha=0.05,details=TRUE)
results_lsmeans
```

The analysis identified three distinct tiers of treatment effectiveness on emergence. Treatment 2 (mean = 55.6) had significantly lower emergence than all others. Treatment 7 (mean = 188.0) showed moderate emergence, higher than Treatment 2 but statistically similar to Treatments 1, 9, 4, and 8 (means = 190.2–197.9). The highest emergence occurred in Treatments 6, 3, and 5 (means = 199.0–200.9), which were statistically similar from one another and from the Treatment 1,9,4 and 8. While Treatments 6, 3, and 5 outperformed Treatment 7 (p < 0.05), there were no significant differences within the upper group or between the upper and intermediate clusters, suggesting most treatments (except Treatment 2) perform similarly.

# Question 5

The provided function lets you dynamically add a linear model plus one factor from that model and plots a bar chart with letters denoting treatment differences. Use this model to generate the plot shown below. Explain the significance of the letters. 
```{r}
plot_cldbars_onefactor <- function(lm_model, factor) {
  data <- lm_model$model
  variables <- colnames(lm_model$model)
  dependent_var <- variables[1]
  independent_var <- variables[2:length(variables)]

  lsmeans <- emmeans(lm_model, as.formula(paste("~", factor))) # estimate lsmeans 
  Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE, Letters = letters) # contrast with Tukey adjustment by default.
  
  # Extracting the letters for the bars
  sig.diff.letters <- data.frame(Results_lsmeans$emmeans[,1], 
                                 str_trim(Results_lsmeans$emmeans[,7]))
  colnames(sig.diff.letters) <- c(factor, "Letters")
  
  # for plotting with letters from significance test
  ave_stand2 <- lm_model$model %>%
    group_by(!!sym(factor)) %>%
    dplyr::summarize(
      ave.emerge = mean(.data[[dependent_var]], na.rm = TRUE),
      se = sd(.data[[dependent_var]]) / sqrt(n())
    ) %>%
    left_join(sig.diff.letters, by = factor) %>%
    mutate(letter_position = ave.emerge + 10 * se)
  
  plot <- ggplot(data, aes(x = !! sym(factor), y = !! sym(dependent_var))) + 
    stat_summary(fun = mean, geom = "bar") +
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +
    ylab("Number of emerged plants") + 
    geom_jitter(width = 0.02, alpha = 0.5) +
    geom_text(data = ave_stand2, aes(label = Letters, y = letter_position), size = 5) +
    xlab(as.character(factor)) +
    theme_classic()
  
  return(plot)
}


plot_cldbars_onefactor(lm2,"Treatment")

```

The letters above each bar represent statistical groupings from a Tukey HSD test (α = 0.05). Treatments sharing the same letter (e.g., "a", "ab") are not significantly different in emergence. For example, if Treatment 2 is labeled "c" (lowest bar), it is significantly worse than all others. Treatment 7, the second lowest, (labelled "b") is statistically similar to treatments 1, 4, 8, and 9 that are labelled with overlapping letters ( ab ). These treatment labels also overlap with "a" indicating no significant differences with higher treatments ( 3, 5, 6 labeled "a").

