---
title: "Homework_LM"
author: "Arpan"
date: "2025-04-01"
output:
  html_document:
    toc: true
    toc_float: true
  md_document:
    variant: gfm
---
# Linear model
Lets load in our mtcars dataset to demonstrate some basic linear regression.

```{r Obtaining data}
data("mtcars") #loading the biultin database
```

```{r Using ggplot}
library(ggplot2)
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_smooth(method = lm, se = FALSE, color = "grey") +
  geom_point(aes(color = wt)) +
  xlab("Weight") + 
  ylab("Miles per gallon") +
  scale_colour_gradient(low = "forestgreen", high = "red") +
  theme_classic()
```
Looks like there is a relationship? How do we know? 

We run a linear model - or if its a continuous x variable and a continuous y variable we would call it a regression. If we are looking to call it a cause and effect relationship we may call it a correlation. 

The simplest way to do this is with the function `lm()` and run summary of this linear model to output some summary statistics
```{r running linear model}
lm1<-lm(mpg~wt,data=mtcars)
summary(lm1)
```

We could run an ANOVA. 
```{r}
anova(lm1)
```
We can also run correlation analysis
```{r}
cor.test(mtcars$wt, mtcars$mpg)
```
This gives us a different r value which is the correlation statistic. 
The closer to -1 or 1 means the stronger the correlation between the two variables. 

**The pvalue is the same in each of these analysis**

## Using a dataset for linear model analyses
```{r}
bull.rich <- read.csv("LinearModel/Bull_richness.csv")
```

Filter our dataset to include one treatment and growth stage for demonstration of a t-test.
```{r}
library(tidyverse)
bull.rich.sub<-bull.rich %>%
  filter(GrowthStage == "V8" & Treatment == "Conv.")
```

```{r}
t.test(richness~Fungicide, data = bull.rich.sub)
```
We can use lm() or anova to get same result.
```{r}
summary(lm(richness~Fungicide, data = bull.rich.sub))
anova(lm(richness~Fungicide, data = bull.rich.sub))
```
In reality, they are all just linear models

## Linear model in Categorical variable
What if our variable is Categorical?
Lets filter our dataset to have only richness in different crop growth stages in the control samples in soybean in conventional management. Here we have three growth stages. 
```{r}
bull.rich.sub2 <- bull.rich %>%
  filter(Fungicide == "C" & Treatment == "Conv." & Crop == "Corn")
```
Lets visualize our data
```{r}
ggplot(bull.rich.sub2, aes(x = GrowthStage, y = richness)) +
  geom_boxplot()
```

```{r}
summary(lm(richness ~ GrowthStage, data = bull.rich.sub2))
anova(lm(richness ~ GrowthStage, data = bull.rich.sub2))
```

## emmeans

But now how do we know which is different from the the other group?

We can do post-hoc tests to find out. The simplest way to think of this is individual t-tests across groups. The most versitile way to do this is with the packages `emmeans`, and `multcomp`. 

The lsmeans are the least squared means - they are the means estimated by the linear model. This is in contrast to the arethmatic means which are the means calculated or the average.
```{r}
library(emmeans)
library(multcomp)
library(multcompView)
lm3<-lm(richness ~ GrowthStage, data = bull.rich.sub2)
lsmeans <- emmeans(lm3, ~GrowthStage) # estimate lsmeans of variety within siteXyear
results_lsmeans <- cld(lsmeans, alpha = 0.05, details = TRUE) # contrast with Tukey ajustment by default. 
results_lsmeans
```

# Interaction
What if we care about the interactions between factors?

Lets filter our dataset to include fungicide term. 

```{r}
bull.rich.sub3 <- bull.rich %>%
  filter(Treatment == "Conv." & Crop == "Corn")

bull.rich.sub3$GrowthStage <- factor(bull.rich.sub3$GrowthStage, levels = c("V6", "V8", "V15"))
```
Now lets set up our linear model with fungicide interaction factor
```{r}
# write it like this
lm.interaction <- lm(richness ~ GrowthStage * Fungicide, data = bull.rich.sub3)# using * gives interaction 
summary(lm.interaction)
anova(lm.interaction)
```

Conducting post-hoc test within the levels of growthstage. 
```{r}
lsmeans<-emmeans(lm.interaction,~Fungicide|GrowthStage)#Growthstages levels within fungicide 
results_lsmeanss<-cld(lsmeans,alpha=0.05,details=TRUE)
results_lsmeanss
```

By visualization we can see that it looks like V8 fungicide affected richness more than others and by V15 they recovered.

```{r}
ggplot(bull.rich.sub3,aes(x=GrowthStage,y=richness, colour=Fungicide))+
  geom_boxplot()
```

# Mixed effects models
In mixed effects models we have fixed and random effects term. The random effects term is something that affects the variation in y. A fixed effect is something that affects the mean of y.
```{r}
library(lme4)
lm.interaction2<-lmer(richness~GrowthStage*Fungicide+(1|Rep),data=bull.rich.sub3)
summary(lm.interaction)
summary(lm.interaction2)

lsmeanss<-emmeans(lm.interaction2,~Fungicide|GrowthStage)
results_lsmeanss2<-cld(lsmeanss,alpha=0.05,details=TRUE)
results_lsmeanss2
```

